---
title: "Preferences Free-Choice (Study 2)"
author: "Bartosz Maćkiewicz"
format:
  html:
    theme: readable
    toc: true
    toc-location: left
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(readr)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(flextable)
library(scales)
library(glue)
```

# Overview

- Task: truth-value judgments (`B` = TRUE, `N` = FALSE) for segmented sentences about food preferences.
- Design: 32 practice trials (feedback) + 114 main trials (no feedback; break halfway).
- Main conditions in the experimental session:
  - Controls: `SINGLE_CONTROL`, `DOUBLE_CONTROL`, `PREFER_DIFFERENT`, `PREFER_SAME` (known truth values).
  - Critical: `EXPERIMENTAL` (“wants … A or B” with `wants|accepts` vs `wants|dislikes`).

# Data Preparation

```{r data_preparation}
data_raw <- read_csv("results.csv", show_col_types = FALSE)

data_all <- data_raw %>%
  mutate(
    resp = case_when(
      response %in% c("b", "B") ~ 1,
      response %in% c("n", "N") ~ 0,
      TRUE ~ NA_real_
    ),
    trial_type = factor(case_when(
      (item1_type == "accepts" & item2_type == "wants") | (item1_type == "wants" & item2_type == "accepts") ~ "Wants + Accepts",
      (item1_type == "accepts" & item2_type == "dislikes") | (item1_type == "dislikes" & item2_type == "accepts") ~ "Dislikes + Accepts",
      (item1_type == "wants" & item2_type == "dislikes") | (item1_type == "dislikes" & item2_type == "wants") ~ "Wants + Dislikes",
      (item1_type == "wants" & item2_type == "wants") ~ "Wants + Wants",
      (item1_type == "dislikes" & item2_type == "dislikes") ~ "Dislikes + Dislikes",
      (item1_type == "accepts" & item2_type == "accepts") ~ "Accepts + Accepts",
      (item1_type == "wants" & (is.na(item2_type) | item2_type == "")) ~ "Wants",
      (item1_type == "dislikes" & (is.na(item2_type) | item2_type == "")) ~ "Dislikes",
      (item1_type == "accepts" & (is.na(item2_type) | item2_type == "")) ~ "Accepts"
    )),
    condition_label = factor(
      case_when(
        condition == "DOUBLE_CONTROL" ~ "Double Control",
        condition == "SINGLE_CONTROL" ~ "Single Control",
        condition == "PREFER_DIFFERENT" ~ "Prefer Different",
        condition == "PREFER_SAME" ~ "Prefer Same",
        condition == "EXPERIMENTAL" ~ "Experimental"
      ),
      levels = c("Double Control", "Single Control", "Prefer Different", "Prefer Same", "Experimental")
    ),
    condition_panel = factor(
      case_when(
        condition %in% c("PREFER_DIFFERENT", "PREFER_SAME") ~ "Prefer",
        condition == "DOUBLE_CONTROL" ~ "Double Control",
        condition == "SINGLE_CONTROL" ~ "Single Control",
        condition == "EXPERIMENTAL" ~ "Experimental"
      ),
      levels = c("Double Control", "Single Control", "Prefer", "Experimental")
    ),
    resp_adj = case_when(
      correct == "Y" & resp == 1 ~ 1,
      correct == "N" & resp == 0 ~ 1,
      correct == "Y/N" & resp == 1 ~ 1,
      TRUE ~ 0
    ),
    participant = factor(PARTICIPANT),
    item_id = factor(item),
    tribe = factor(tribe, levels = c("Urbanite", "Nomad")),
    rt_sec = rt / 1000,
    log_rt = log(rt_sec)
  )

data_practice <- data_all %>% filter(SESSION == "PRACTICE")
data_main <- data_all %>% filter(SESSION == "EXPERIMENTAL")

glue(
  "Loaded {n_distinct(data_main$participant)} participant(s) with {nrow(data_practice)} practice trials and {nrow(data_main)} main-session trials."
)
```

# Stimuli inventory (paper methods)

```{r stimuli_inventory, include=FALSE}
practice_stimuli <- data_practice %>%
  distinct(
    item, condition, condition_label, verb, trial_type, correct, structure,
    subcondition, order, relation, tribe, item1_type, item2_type
  )

main_stimuli <- data_main %>%
  distinct(
    item, condition, condition_label, verb, trial_type, correct, structure,
    subcondition, order, relation, tribe, item1_type, item2_type
  )

stimuli_inventory_text <- glue(
  "Unique stimuli counts: practice = {nrow(practice_stimuli)} item(s), main session = {nrow(main_stimuli)} item(s)."
)

main_counts_by_condition <- main_stimuli %>%
  count(condition_label, name = "N items") %>%
  arrange(condition_label)

main_counts_by_condition_and_type <- main_stimuli %>%
  count(condition_label, trial_type, correct, name = "N items") %>%
  arrange(condition_label, trial_type, correct)

critical_counts <- main_stimuli %>%
  filter(condition == "EXPERIMENTAL", verb == "want") %>%
  mutate(
    other_type = if_else(item1_type == "wants", item2_type, item1_type),
    other_type = factor(other_type, levels = c("accepts", "dislikes")),
    wants_first = item1_type == "wants"
  ) %>%
  count(other_type, wants_first, name = "N items") %>%
  arrange(other_type, desc(wants_first))

NULL
```

`r stimuli_inventory_text`

Main-session items by condition (unique stimuli):

```{r stimuli_inventory_main_by_condition, results='asis'}
main_counts_by_condition %>% flextable() %>% autofit()
```

Main-session items by condition and item type (unique stimuli):

```{r stimuli_inventory_main_detail, results='asis'}
main_counts_by_condition_and_type %>% flextable() %>% autofit()
```

Critical (EXPERIMENTAL) items by subcondition and order (unique stimuli):

```{r stimuli_inventory_critical, results='asis'}
critical_counts %>%
  mutate(
    other_type = as.character(other_type),
    wants_first = if_else(wants_first, "wants-first", "wants-second")
  ) %>%
  rename(
    `Non-wanted option` = other_type,
    `Order` = wants_first
  ) %>%
  flextable() %>%
  autofit()
```

# Plotting Theme and Palettes

```{r theme_and_palettes}
pal_trial_type <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999", "#000000"
)

theme_report <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(colour = "grey40"),
      strip.background = element_rect(fill = "grey95", colour = "grey80"),
      strip.text = element_text(face = "bold", size = rel(0.7)),
      legend.title = element_text(face = "bold"),
      plot.title = element_text(face = "bold", size = rel(1.2)),
      plot.subtitle = element_text(size = rel(0.9), color = "grey30"),
      plot.caption = element_text(hjust = 0, size = rel(0.8), color = "grey50")
    )
}

theme_set(theme_report())
```

# Demographics (Prolific export)

```{r}
prolific_files <- list.files(".", pattern = "^prolific_export_.*\\.csv$", full.names = TRUE)
if (length(prolific_files) == 0) stop("No Prolific export CSV found in the results folder.")
prolific_path <- prolific_files[which.max(file.info(prolific_files)$mtime)]

prolific <- read_csv(prolific_path, show_col_types = FALSE) %>%
  rename(
    PROLIFIC_PID = `Participant id`,
    Submission_ID = `Submission id`,
    Status = Status,
    Time_taken_sec = `Time taken`,
    Age_years = Age,
    Sex = Sex,
    Ethnicity = `Ethnicity simplified`,
    Country_birth = `Country of birth`,
    Country_residence = `Country of residence`,
    Nationality = Nationality,
    Language = Language,
    Student_status = `Student status`,
    Employment_status = `Employment status`
  ) %>%
  mutate(Age_years = suppressWarnings(as.numeric(Age_years)))

matched_ids <- unique(data_main$PARTICIPANT)
demog <- prolific %>% filter(PROLIFIC_PID %in% matched_ids)
missing_demog <- setdiff(matched_ids, demog$PROLIFIC_PID)

demog_summary <- demog %>%
  summarise(
    Matched = n_distinct(PROLIFIC_PID),
    Mean_age = mean(Age_years, na.rm = TRUE),
    SD_age = sd(Age_years, na.rm = TRUE),
    Min_age = min(Age_years, na.rm = TRUE),
    Max_age = max(Age_years, na.rm = TRUE)
  )

list(
  prolific_path = prolific_path,
  demog_summary = demog_summary,
  missing_demog = missing_demog
)
```

```{r}
demog_summary %>% flextable() %>% autofit()
```

```{r}
demog %>%
  count(Sex, name = "Count") %>%
  mutate(Percent = percent(Count / sum(Count))) %>%
  flextable() %>%
  autofit()
```

```{r}
if (length(missing_demog) > 0) {
  tibble(`Unmatched participant ID` = missing_demog) %>% flextable() %>% autofit()
} else {
  cat("All participants with data matched to Prolific demographics.")
}
```

# Paper-ready summaries

## Descriptive TRUE-response rates by item type (main session)

```{r paper_descriptives, include=FALSE}
paper_condition_levels <- c(
  "Single control: wants (true)",
  "Single control: accepts (false)",
  "Single control: dislikes (false)",
  "Double control: wants ∨ wants (true)",
  "Double control: accepts ∨ accepts (false)",
  "Double control: dislikes ∨ dislikes (false)",
  "Prefer different: wants > accepts (true)",
  "Prefer different: wants < accepts (false)",
  "Prefer different: accepts > dislikes (true)",
  "Prefer different: accepts < dislikes (false)",
  "Prefer same: wants = wants (false)",
  "Prefer same: accepts = accepts (false)",
  "Experimental: wants ∨ accepts (wants-first)",
  "Experimental: wants ∨ accepts (wants-second)",
  "Experimental: wants ∨ dislikes (wants-first)",
  "Experimental: wants ∨ dislikes (wants-second)"
)

paper_conditions <- data_main %>%
  mutate(
    other_type = if_else(item1_type == "wants", item2_type, item1_type),
    wants_first = item1_type == "wants",
    paper_condition = case_when(
      condition == "SINGLE_CONTROL" & item1_type == "wants" ~ "Single control: wants (true)",
      condition == "SINGLE_CONTROL" & item1_type == "accepts" ~ "Single control: accepts (false)",
      condition == "SINGLE_CONTROL" & item1_type == "dislikes" ~ "Single control: dislikes (false)",
      condition == "DOUBLE_CONTROL" & trial_type == "Wants + Wants" ~ "Double control: wants ∨ wants (true)",
      condition == "DOUBLE_CONTROL" & trial_type == "Accepts + Accepts" ~ "Double control: accepts ∨ accepts (false)",
      condition == "DOUBLE_CONTROL" & trial_type == "Dislikes + Dislikes" ~ "Double control: dislikes ∨ dislikes (false)",
      condition == "PREFER_DIFFERENT" & trial_type == "Wants + Accepts" & correct == "Y" ~ "Prefer different: wants > accepts (true)",
      condition == "PREFER_DIFFERENT" & trial_type == "Wants + Accepts" & correct == "N" ~ "Prefer different: wants < accepts (false)",
      condition == "PREFER_DIFFERENT" & trial_type == "Dislikes + Accepts" & correct == "Y" ~ "Prefer different: accepts > dislikes (true)",
      condition == "PREFER_DIFFERENT" & trial_type == "Dislikes + Accepts" & correct == "N" ~ "Prefer different: accepts < dislikes (false)",
      condition == "PREFER_SAME" & trial_type == "Wants + Wants" ~ "Prefer same: wants = wants (false)",
      condition == "PREFER_SAME" & trial_type == "Accepts + Accepts" ~ "Prefer same: accepts = accepts (false)",
      condition == "EXPERIMENTAL" & other_type == "accepts" & wants_first ~ "Experimental: wants ∨ accepts (wants-first)",
      condition == "EXPERIMENTAL" & other_type == "accepts" & !wants_first ~ "Experimental: wants ∨ accepts (wants-second)",
      condition == "EXPERIMENTAL" & other_type == "dislikes" & wants_first ~ "Experimental: wants ∨ dislikes (wants-first)",
      condition == "EXPERIMENTAL" & other_type == "dislikes" & !wants_first ~ "Experimental: wants ∨ dislikes (wants-second)",
      TRUE ~ NA_character_
    ),
    paper_condition = factor(paper_condition, levels = paper_condition_levels)
  ) %>%
  filter(!is.na(paper_condition)) %>%
  group_by(participant, paper_condition) %>%
  summarise(mean_true = mean(resp, na.rm = TRUE), .groups = "drop")

paper_descriptives <- paper_conditions %>%
  group_by(paper_condition) %>%
  summarise(
    n_participants = n_distinct(participant),
    mean = mean(mean_true, na.rm = TRUE),
    sd = sd(mean_true, na.rm = TRUE),
    se = sd / sqrt(n_participants),
    .groups = "drop"
  ) %>%
  mutate(
    ci_mult = qt(0.975, df = n_participants - 1),
    ci_low = mean - ci_mult * se,
    ci_high = mean + ci_mult * se,
    `Mean TRUE` = percent(mean, accuracy = 0.1),
    `SD` = percent(sd, accuracy = 0.1),
    `95% CI` = paste0("[", percent(ci_low, accuracy = 0.1), ", ", percent(ci_high, accuracy = 0.1), "]")
  ) %>%
  select(
    Condition = paper_condition,
    `Mean TRUE`,
    SD,
    `95% CI`,
    `N participants` = n_participants
  )

NULL
```

```{r paper_descriptives_table, results='asis'}
paper_descriptives %>% flextable() %>% autofit()
```

```{r paper_descriptives_plot}
paper_plot_data <- paper_conditions %>%
  group_by(paper_condition) %>%
  summarise(
    mean = mean(mean_true, na.rm = TRUE),
    se = sd(mean_true, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(paper_plot_data, aes(paper_condition, mean)) +
  geom_col(fill = "#5DADE2", width = 0.75) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "TRUE-response rates by item type (main session)",
    subtitle = "Participant means; error bars are SEM across participants",
    x = NULL,
    y = "Mean TRUE-response rate"
  )
```

# Part 1: Response Analysis

### Proportion of "TRUE" Responses (Raw Data)

```{r raw_response_plot}
subject_means_raw <- data_main %>%
  group_by(PARTICIPANT, condition_panel, verb, trial_type, correct) %>%
  summarise(mean_resp = mean(resp, na.rm = TRUE), .groups = "drop")

summary_means_raw <- subject_means_raw %>%
  group_by(condition_panel, verb, trial_type, correct) %>%
  summarise(
    mean = mean(mean_resp),
    se = sd(mean_resp) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_means_raw, aes(x = correct, y = mean, fill = trial_type)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.8),
    width = 0.2,
    linewidth = 0.6
  ) +
  facet_wrap(~condition_panel, ncol = 2) +
  scale_fill_manual(values = pal_trial_type, name = "Trial Type") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Percentage of 'TRUE' Responses",
    subtitle = "Participant means; error bars are standard errors.",
    x = "Correctness Category",
    y = "Mean 'TRUE' Response Rate"
  )
```

### Proportion of "TRUE" Responses (Raw Data) (better plots)

```{r raw_response_plot_ghosted}
ggplot(summary_means_raw, aes(x = correct, y = mean, fill = trial_type, alpha = condition_panel == "Experimental")) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.8),
    width = 0.2,
    linewidth = 0.6
  ) +
  facet_wrap(~condition_panel, ncol = 2) +
  scale_fill_manual(values = pal_trial_type, name = "Trial Type") +
  scale_alpha_manual(values = c(`TRUE` = 1, `FALSE` = 0.25), guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Percentage of 'TRUE' Responses",
    subtitle = "Experimental highlighted; other conditions ghosted",
    x = "Correctness Category",
    y = "Mean 'TRUE' Response Rate"
  )
```

### Want Controls (Sanity Check)

```{r}
want_controls <- data_main %>%
  filter(condition %in% c("SINGLE_CONTROL", "DOUBLE_CONTROL")) %>%
  filter(verb == "want") %>%
  mutate(
    item1_type = factor(item1_type, levels = c("wants", "accepts", "dislikes"))
  )

want_summary <- want_controls %>%
  group_by(condition, item1_type, correct) %>%
  summarise(
    n = n(),
    true_rate = mean(resp, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(condition, item1_type)

want_summary %>%
  mutate(
    true_rate = percent(true_rate, accuracy = 0.1)
  ) %>%
  flextable() %>%
  autofit()
```

```{r}
want_subj <- want_controls %>%
  group_by(participant, condition, item1_type) %>%
  summarise(mean_true = mean(resp, na.rm = TRUE), .groups = "drop")

want_plot <- want_subj %>%
  group_by(condition, item1_type) %>%
  summarise(
    mean = mean(mean_true),
    se = sd(mean_true) / sqrt(n()),
    .groups = "drop"
  )

ggplot(want_plot, aes(x = item1_type, y = mean, fill = item1_type)) +
  geom_col(width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  facet_wrap(~ condition, nrow = 1) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_fill_manual(values = c(wants = "#2C7FB8", accepts = "#FDAE61", dislikes = "#D7191C"), guide = "none") +
  labs(
    title = "TRUE-response rate on 'want' controls by item type",
    subtitle = "Error bars = SEM over participant means",
    x = NULL,
    y = NULL
  )
```

## Distribution of Responses in Experimental Condition

```{r distribution_plot}
prop_by_part <- data_main %>%
  filter(condition == "EXPERIMENTAL", verb == "want") %>%
  group_by(PARTICIPANT, trial_type) %>%
  summarise(p = mean(resp, na.rm = TRUE), .groups = "drop")

ggplot(prop_by_part, aes(x = p)) +
  geom_histogram(bins = 12, fill = "#5DADE2", color = "white", alpha = 0.9) +
  geom_vline(xintercept = c(.10, .90), linetype = "dashed", color = "firebrick") +
  scale_x_continuous(labels = percent_format()) +
  facet_wrap(~trial_type) +
  labs(
    title = "Distribution of 'TRUE' Response Rates per Participant",
    subtitle = "Condition: EXPERIMENTAL (want). Dashed lines at 10% and 90%.",
    x = "Participant Mean TRUE-response Rate",
    y = "Number of Participants"
  )
```

### Experimental Condition (Critical Trials)

```{r}
critical <- data_main %>%
  filter(condition == "EXPERIMENTAL", verb == "want") %>%
  mutate(
    other_type = if_else(item1_type == "wants", item2_type, item1_type),
    other_type = factor(other_type, levels = c("accepts", "dislikes")),
    wants_first = item1_type == "wants"
  )

subj_means <- critical %>%
  group_by(participant, other_type, wants_first) %>%
  summarise(mean_true = mean(resp, na.rm = TRUE), .groups = "drop")

summary_means <- subj_means %>%
  group_by(other_type, wants_first) %>%
  summarise(
    mean = mean(mean_true),
    se = sd(mean_true) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_means, aes(x = other_type, y = mean, fill = wants_first)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.7),
    width = 0.15,
    linewidth = 0.6
  ) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_fill_manual(values = c("TRUE" = "#2C7FB8", "FALSE" = "#F03B20")) +
  labs(
    title = "Proportion of TRUE responses on critical trials",
    subtitle = "Error bars = SEM over participant means",
    x = "Non-wanted option type",
    y = NULL,
    fill = "Wanted option first"
  )
```

### Mixed-Effects Model (Critical Trials)

```{r}
critical_model <- glmer(
  resp ~ other_type * wants_first + tribe + (1 | participant) + (1 | item_id),
  data = critical,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
tab_model(critical_model)
```

```{r}
emm <- emmeans(critical_model, ~ other_type * wants_first, type = "response")
as.data.frame(summary(emm, infer = c(TRUE, TRUE))) %>%
  mutate(across(c(prob, asymp.LCL, asymp.UCL), ~ scales::percent(.x, accuracy = 0.1))) %>%
  select(other_type, wants_first, prob, asymp.LCL, asymp.UCL, z.ratio, p.value) %>%
  flextable() %>%
  autofit()
```

#### Midpoint (50%) tests (model-based; emmeans)

```{r}
# Midpoint (50%) tests using the fitted mixed-effects model.
# Note: tests are performed on the logit scale (so 0 corresponds to 50% on the response scale).
emm_midpoint <- emmeans(critical_model, ~ other_type, weights = "proportional")

midpoint_model_tbl <- as.data.frame(
  summary(emm_midpoint, type = "response", infer = c(TRUE, TRUE))
) %>%
  mutate(
    `Disjunction type` = if_else(
      other_type == "accepts",
      "wants ∨ accepts (W+A)",
      "wants ∨ dislikes (W+D)"
    ),
    Direction = if_else(prob > 0.5, "above 50%", "below 50%"),
    `Predicted TRUE` = percent(prob, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      percent(asymp.LCL, accuracy = 0.1),
      ", ",
      percent(asymp.UCL, accuracy = 0.1),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(`Disjunction type`, Direction, `Predicted TRUE`, `95% CI`, z, p)

midpoint_model_tbl %>% flextable() %>% autofit()
```

#### Midpoint (50%) tests (participant means; descriptive)

```{r}
# Midpoint (50%) tests on participant-mean TRUE-response rates, collapsed over disjunct order.
subj_means_by_other <- critical %>%
  group_by(participant, other_type) %>%
  summarise(mean_true = mean(resp, na.rm = TRUE), .groups = "drop")

midpoint_tests_tbl <- subj_means_by_other %>%
  group_by(other_type) %>%
  group_modify(~{
    tt <- t.test(.x$mean_true, mu = 0.5)
    tibble(
      N = nrow(.x),
      Mean = mean(.x$mean_true),
      SD = sd(.x$mean_true),
      CI_low = tt$conf.int[1],
      CI_high = tt$conf.int[2],
      t = unname(tt$statistic),
      df = unname(tt$parameter),
      p = tt$p.value
    )
  }) %>%
  ungroup() %>%
  mutate(
    `Disjunction type` = if_else(
      other_type == "accepts",
      "wants ∨ accepts (W+A)",
      "wants ∨ dislikes (W+D)"
    ),
    Direction = if_else(Mean > 0.5, "above 50%", "below 50%"),
    Mean = percent(Mean, accuracy = 0.1),
    SD = percent(SD, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      percent(CI_low, accuracy = 0.1),
      ", ",
      percent(CI_high, accuracy = 0.1),
      "]"
    ),
    t = round(t, 2),
    df = round(df, 0),
    p = format.pval(p, digits = 3, eps = .001)
  ) %>%
  select(`Disjunction type`, N, Direction, Mean, SD, `95% CI`, t, df, p)

midpoint_tests_tbl %>% flextable() %>% autofit()
```

```{r}
emm_other_by_order <- emmeans(critical_model, ~ other_type | wants_first)
other_contrasts <- summary(pairs(emm_other_by_order), infer = c(TRUE, TRUE))

other_contrasts_tbl <- as.data.frame(other_contrasts) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    wants_first = if_else(wants_first, "wants-first", "wants-second"),
    `OR (95% CI)` = paste0(
      round(OR, 2),
      " [",
      round(OR_low, 2),
      ", ",
      round(OR_high, 2),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(
    `Order` = wants_first,
    Contrast = contrast,
    `OR (95% CI)`,
    z,
    p
  )

other_contrasts_tbl %>% flextable() %>% autofit()
```

```{r}
emm_order_by_other <- emmeans(critical_model, ~ wants_first | other_type)
order_contrasts <- summary(pairs(emm_order_by_other), infer = c(TRUE, TRUE))

order_contrasts_tbl <- as.data.frame(order_contrasts) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(
      round(OR, 2),
      " [",
      round(OR_low, 2),
      ", ",
      round(OR_high, 2),
      "]"
    ),
    contrast = if_else(contrast == "FALSE - TRUE", "wants-second - wants-first", contrast),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(
    `Non-wanted option` = other_type,
    Contrast = contrast,
    `OR (95% CI)`,
    z,
    p
  )

order_contrasts_tbl %>% flextable() %>% autofit()
```

# Part 2: Response Time (RT) Analysis

```{r rt_cleaning}
filter_rt_outliers <- function(data, rt_col = "rt", lower_ms = 200, mad_cutoff = 2.5) {
  data %>%
    filter(.data[[rt_col]] >= lower_ms) %>%
    mutate(log_rt = log(.data[[rt_col]])) %>%
    group_by(PARTICIPANT) %>%
    mutate(
      med_log = median(log_rt, na.rm = TRUE),
      mad_log = mad(log_rt, constant = 1, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    filter(abs(log_rt - med_log) <= mad_cutoff * mad_log) %>%
    select(-log_rt, -med_log, -mad_log)
}

critical_rt <- critical %>%
  filter(!is.na(rt)) %>%
  filter_rt_outliers() %>%
  mutate(rt_log = log(rt))
```

### Mixed-Effects Model (Log RT; Critical Trials)

```{r}
rt_model <- lmer(
  rt_log ~ other_type * wants_first + tribe + (1 | participant) + (1 | item_id),
  data = critical_rt
)

tab_model(rt_model)
```
```
