---
title: "Preferences Free-Choice (Study 2): Critical want disjunctions"
author: "Bartosz"
output:
  html_document:
    theme: readable
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(readr)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(flextable)
library(scales)
library(glue)
```

# Data preparation

```{r data_preparation}
data_raw <- read_csv("results.csv", show_col_types = FALSE)

data_main <- data_raw %>%
  mutate(
    resp = case_when(
      response %in% c("b", "B") ~ 1,
      response %in% c("n", "N") ~ 0,
      TRUE ~ NA_real_
    ),
    trial_type = factor(case_when(
      (item1_type == "accepts" & item2_type == "wants") | (item1_type == "wants" & item2_type == "accepts") ~ "Wants + Accepts",
      (item1_type == "accepts" & item2_type == "dislikes") | (item1_type == "dislikes" & item2_type == "accepts") ~ "Dislikes + Accepts",
      (item1_type == "wants" & item2_type == "dislikes") | (item1_type == "dislikes" & item2_type == "wants") ~ "Wants + Dislikes",
      (item1_type == "wants" & item2_type == "wants") ~ "Wants + Wants",
      (item1_type == "dislikes" & item2_type == "dislikes") ~ "Dislikes + Dislikes",
      (item1_type == "accepts" & item2_type == "accepts") ~ "Accepts + Accepts",
      (item1_type == "wants" & (is.na(item2_type) | item2_type == "")) ~ "Wants",
      (item1_type == "dislikes" & (is.na(item2_type) | item2_type == "")) ~ "Dislikes",
      (item1_type == "accepts" & (is.na(item2_type) | item2_type == "")) ~ "Accepts"
    )),
    participant = factor(PARTICIPANT),
    item_id = factor(item),
    tribe = factor(tribe, levels = c("Urbanite", "Nomad"))
  ) %>%
  filter(SESSION == "EXPERIMENTAL")

glue(
  "Loaded {n_distinct(data_main$participant)} participant(s) and {nrow(data_main)} main-session trial(s)."
)
```

# Want disjunctions (order-collapsed): single overarching model

```{r critical_subset}
want_disj <- data_main %>%
  filter(
    verb == "want",
    trial_type %in% c(
      "Wants + Wants",
      "Accepts + Accepts",
      "Dislikes + Dislikes",
      "Wants + Accepts",
      "Wants + Dislikes"
    )
  ) %>%
  mutate(
    disj_type = case_when(
      trial_type == "Wants + Wants" ~ "W+W",
      trial_type == "Wants + Accepts" ~ "W+A",
      trial_type == "Wants + Dislikes" ~ "W+D",
      trial_type == "Accepts + Accepts" ~ "A+A",
      trial_type == "Dislikes + Dislikes" ~ "D+D",
      TRUE ~ NA_character_
    ),
    disj_type = factor(disj_type, levels = c("W+W", "W+A", "W+D", "A+A", "D+D"))
  ) %>%
  filter(!is.na(resp), !is.na(disj_type))

glue(
  "Want-disjunction trials: {nrow(want_disj)}; participants: {n_distinct(want_disj$participant)}; items: {n_distinct(want_disj$item_id)}."
)
```

## Mixed-effects model (want disjunctions)

```{r critical_model}
critical_model <- glmer(
  resp ~ disj_type + tribe + (1 | participant) + (1 | item_id),
  data = want_disj,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

tab_model(critical_model)
```

## Model-based estimates (emmeans)

```{r critical_emmeans}
emm <- emmeans(critical_model, ~ disj_type, weights = "proportional", type = "response")

as.data.frame(summary(emm, infer = c(TRUE, TRUE))) %>%
  mutate(across(c(prob, asymp.LCL, asymp.UCL), ~ scales::percent(.x, accuracy = 0.1))) %>%
  select(disj_type, prob, asymp.LCL, asymp.UCL, z.ratio, p.value) %>%
  flextable() %>%
  autofit()
```

## W+A vs W+D (order-collapsed; model-based)

```{r critical_overall_contrast}
emm_disj <- emmeans(critical_model, ~ disj_type, weights = "proportional")

overall_contrast <- summary(
  contrast(
    emm_disj,
    method = list(
      "W+A - W+D" = c(0, 1, -1, 0, 0)
    )
  ),
  infer = c(TRUE, TRUE)
)

as.data.frame(overall_contrast) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(
      round(OR, 2),
      " [",
      round(OR_low, 2),
      ", ",
      round(OR_high, 2),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(contrast, `OR (95% CI)`, z, p) %>%
  flextable() %>%
  autofit()
```

## Midpoint (50%) tests (model-based; emmeans)

```{r critical_midpoint_emmeans}
# Midpoint tests on model-estimated marginal probabilities (averaged over tribe).
# Note: tests are performed on the logit scale (so 0 corresponds to 50% on the response scale).
emm_midpoint <- emmeans(critical_model, ~ disj_type, weights = "proportional")

midpoint_model_tbl <- as.data.frame(summary(emm_midpoint, type = "response", infer = c(TRUE, TRUE))) %>%
  mutate(
    Direction = if_else(prob > 0.5, "above 50%", "below 50%"),
    `Predicted TRUE` = percent(prob, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      percent(asymp.LCL, accuracy = 0.1),
      ", ",
      percent(asymp.UCL, accuracy = 0.1),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  filter(disj_type %in% c("W+A", "W+D")) %>%
  mutate(`Disjunction type` = recode(disj_type, `W+A` = "wants ∨ accepts (W+A)", `W+D` = "wants ∨ dislikes (W+D)")) %>%
  select(`Disjunction type`, Direction, `Predicted TRUE`, `95% CI`, z, p)

midpoint_model_tbl %>% flextable() %>% autofit()
```

# Benchmark: W+W vs W+A (order-collapsed; model-based)

```{r ww_vs_wa_contrast}
emm_disj <- emmeans(critical_model, ~ disj_type, weights = "proportional")

ww_vs_wa_contrast <- summary(
  contrast(
    emm_disj,
    method = list(
      "W+A - W+W" = c(-1, 1, 0, 0, 0)
    )
  ),
  infer = c(TRUE, TRUE)
)

as.data.frame(ww_vs_wa_contrast) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(
      round(OR, 2),
      " [",
      round(OR_low, 2),
      ", ",
      round(OR_high, 2),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(contrast, `OR (95% CI)`, z, p) %>%
  flextable() %>%
  autofit()
```

# Benchmark: W+W vs W+D (order-collapsed; model-based)

```{r ww_vs_wd_contrast}
emm_disj <- emmeans(critical_model, ~ disj_type, weights = "proportional")

ww_vs_wd_contrast <- summary(
  contrast(
    emm_disj,
    method = list(
      "W+D - W+W" = c(-1, 0, 1, 0, 0)
    )
  ),
  infer = c(TRUE, TRUE)
)

as.data.frame(ww_vs_wd_contrast) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(
      round(OR, 2),
      " [",
      round(OR_low, 2),
      ", ",
      round(OR_high, 2),
      "]"
    ),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(contrast, `OR (95% CI)`, z, p) %>%
  flextable() %>%
  autofit()
```

## Tribe effect (Nomad vs Urbanite; model-based)

```{r tribe_effect}
beta <- fixef(critical_model)["tribeNomad"]
se <- sqrt(vcov(critical_model)["tribeNomad", "tribeNomad"])
ci <- beta + c(-1, 1) * 1.96 * se

tribe_tbl <- tibble(
  `Effect` = "Nomad vs Urbanite",
  `OR (95% CI)` = paste0(
    round(exp(beta), 2),
    " [",
    round(exp(ci[1]), 2),
    ", ",
    round(exp(ci[2]), 2),
    "]"
  ),
  z = round(summary(critical_model)$coefficients["tribeNomad", "z value"], 2),
  p = format.pval(summary(critical_model)$coefficients["tribeNomad", "Pr(>|z|)"], digits = 3, eps = .001)
)

tribe_tbl %>% flextable() %>% autofit()
```
