---
title: "Entailment Free-Choice (Study 1)"
author: "Bartosz"
output:
  html_document:
    theme: readable
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(scales)
library(lme4)
library(lmerTest)
library(emmeans)
library(scico)
library(flextable)
library(glue)
library(sjPlot)
library(broom.mixed)
library(writexl)

# cleaner global theme
theme_set(theme_minimal(base_size = 12))
```

# Overview

- Task: judge whether a Conclusion follows from a Premise (keys: `B` = follows, `N` = does not follow).
- Design: 8 practice trials (controls with feedback) + 96 experimental trials (48 critical modal/want; 48 control fillers; no feedback).
- Stimuli metadata fields in `results.csv`: `domain` (`modal`, `want`, `control`), `pair_type` (e.g., `allowed_and`, `want_single`), `conclusion_form` (`A`, `B`, `AND`, or empty for controls), and validity (`STIMULUS_VALID`: `Y`, `N`, or `?` for unknown/critical).

# Data Preparation

```{r data-prep}
data_raw <- read_csv("results.csv", show_col_types = FALSE)

n_participants <- n_distinct(data_raw$PARTICIPANT)
table_session <- table(data_raw$SESSION)

data_main <- data_raw %>%
  mutate(
    resp = case_when(
      response %in% c("b", "B") ~ 1,
      response %in% c("n", "N") ~ 0,
      TRUE ~ NA_real_
    ),
    valid_flag = if_else(STIMULUS_VALID %in% c("Y", "N"), STIMULUS_VALID, NA_character_),
    accuracy = case_when(
      valid_flag == "Y" & resp == 1 ~ 1,
      valid_flag == "N" & resp == 0 ~ 1,
      is.na(valid_flag) ~ NA_real_,
      TRUE ~ 0
    ),
    domain = factor(domain, levels = c("control", "modal", "want")),
    pair_type = fct_inorder(pair_type),
    conclusion_form = fct_inorder(conclusion_form),
    participant = factor(PARTICIPANT),
    rt_sec = rt / 1000,
    log_rt = log(rt_sec)
  )

practice <- filter(data_main, SESSION == "PRACTICE")
experimental <- filter(data_main, SESSION == "EXPERIMENTAL")

data_prep_summary <- glue(
  "Data include {n_participants} participant(s) with {nrow(practice)} practice trials and ",
  "{nrow(experimental)} experimental trials. Session counts: practice = {table_session[['PRACTICE']]}, ",
  "experimental = {table_session[['EXPERIMENTAL']]}. Responses recoded to binary (B=1, N=0) and accuracy ",
  "computed where validity was known."
)
```

Data summary: `r data_prep_summary`

# Stimuli inventory (paper methods)

```{r stimuli-inventory}
unique_practice_items <- practice %>% distinct(item) %>% nrow()
unique_experimental_items <- experimental %>% distinct(item) %>% nrow()

stimuli_experimental <- experimental %>%
  distinct(item, domain, pair_type, conclusion_form, STIMULUS_VALID)

stimuli_practice <- practice %>%
  distinct(item, domain, pair_type, conclusion_form, STIMULUS_VALID)

stimuli_inventory_text <- glue(
  "Unique stimuli counts: practice = {unique_practice_items} item(s), ",
  "experimental = {unique_experimental_items} item(s)."
)

stimuli_counts_by_domain <- stimuli_experimental %>%
  count(domain, name = "N items") %>%
  arrange(domain)

stimuli_counts_practice <- stimuli_practice %>%
  count(pair_type, STIMULUS_VALID, name = "N items") %>%
  arrange(pair_type, STIMULUS_VALID)

stimuli_counts_controls <- stimuli_experimental %>%
  filter(domain == "control") %>%
  count(pair_type, STIMULUS_VALID, name = "N items") %>%
  arrange(pair_type, STIMULUS_VALID)

stimuli_counts_critical <- stimuli_experimental %>%
  filter(domain %in% c("modal", "want")) %>%
  count(domain, pair_type, conclusion_form, name = "N items") %>%
  arrange(domain, pair_type, conclusion_form)

list(
  stimuli_inventory_text = stimuli_inventory_text,
  stimuli_counts_by_domain = stimuli_counts_by_domain,
  stimuli_counts_practice = stimuli_counts_practice,
  stimuli_counts_controls = stimuli_counts_controls,
  stimuli_counts_critical = stimuli_counts_critical
)
```

`r stimuli_inventory_text`

Experimental items by domain (unique stimuli):

```{r stimuli-inventory-domain, results='asis'}
stimuli_counts_by_domain %>% flextable() %>% autofit()
```

Practice items (unique stimuli):

```{r stimuli-inventory-practice, results='asis'}
stimuli_counts_practice %>% flextable() %>% autofit()
```

Experimental control items (unique stimuli):

```{r stimuli-inventory-controls, results='asis'}
stimuli_counts_controls %>% flextable() %>% autofit()
```

Experimental critical items (unique stimuli):

```{r stimuli-inventory-critical, results='asis'}
stimuli_counts_critical %>% flextable() %>% autofit()
```

# Manipulation Checks

```{r manipulation-checks}
trial_counts <- experimental %>%
  count(participant, domain, pair_type) %>%
  arrange(participant, domain, pair_type)

mc_summary <- trial_counts %>%
  group_by(participant) %>%
  summarise(total = sum(n), .groups = "drop") %>%
  summarise(
    min_trials = min(total),
    max_trials = max(total),
    mean_trials = mean(total)
  )

mc_text <- glue(
  "Participants contributed between {mc_summary$min_trials} and {mc_summary$max_trials} experimental trials ",
  "(mean = {round(mc_summary$mean_trials,1)}). Trial counts cover all domains and pair types."
)
```

Manipulation check: `r mc_text`

# Demographics

```{r demographics}
# Locate the Prolific export
prolific_files <- list.files(".", pattern = "^prolific_export_.*\\.csv$", full.names = TRUE)
if (length(prolific_files) == 0) stop("No Prolific export CSV found in the results folder.")
prolific_path <- prolific_files[1]

prolific <- read_csv(prolific_path, show_col_types = FALSE) %>%
  rename(
    PROLIFIC_PID = `Participant id`,
    Age_years = Age,
    Sex = Sex,
    Ethnicity = `Ethnicity simplified`,
    `Country of birth` = `Country of birth`,
    `Country of residence` = `Country of residence`,
    Nationality = Nationality,
    Language = Language,
    `Student status` = `Student status`,
    `Employment status` = `Employment status`
  ) %>%
  mutate(Age_years = suppressWarnings(as.numeric(Age_years)))

matched_ids <- unique(experimental$PARTICIPANT)
demog <- prolific %>% filter(PROLIFIC_PID %in% matched_ids)
missing_demog <- setdiff(matched_ids, demog$PROLIFIC_PID)

demog_summary <- demog %>%
  summarise(
    `Matched participants` = n_distinct(PROLIFIC_PID),
    `Participants in data` = length(matched_ids),
    `Mean age` = mean(Age_years, na.rm = TRUE),
    `SD age` = sd(Age_years, na.rm = TRUE),
    `Min age` = min(Age_years, na.rm = TRUE),
    `Max age` = max(Age_years, na.rm = TRUE)
  )

table_to_flex <- function(df) {
  flextable(df) |> autofit()
}

sex_table <- demog %>% count(Sex, name = "Count") %>%
  mutate(Percent = scales::percent(Count / sum(Count)))
eth_table <- demog %>% count(Ethnicity, name = "Count") %>%
  mutate(Percent = scales::percent(Count / sum(Count)))
country_table <- demog %>% count(`Country of residence`, name = "Count") %>%
  mutate(Percent = scales::percent(Count / sum(Count)))

list(
  demog_summary = demog_summary,
  missing_demog = missing_demog
)
```

Matched demographics summary:

```{r demographics-tables, results='asis'}
table_to_flex(demog_summary)
```

Sex distribution:

```{r demographics-sex, results='asis'}
table_to_flex(sex_table)
```

Ethnicity distribution:

```{r demographics-ethnicity, results='asis'}
table_to_flex(eth_table)
```

Country of residence distribution:

```{r demographics-country, results='asis'}
table_to_flex(country_table)
```

Unmatched participant IDs (if any):

```{r demographics-missing, results='asis'}
if (length(missing_demog) > 0) {
  missing_tbl <- tibble(`Unmatched participant ID` = missing_demog)
  missing_tbl %>% flextable() %>% autofit()
} else {
  cat("All participants with data matched to Prolific demographics.")
}
```

# Response Patterns (Experimental Trials)

### Group definitions

```{r group-definitions}
pair_order <- c(
  "conjunction", "conditional", "quantifier", "presupposition",
  "want_and", "want_mind_and", "want_single", "want_mind_single", "atleast_single_a", "atleast_single_b",
  "allowed_and", "must_and", "allowed_single", "must_single",
  "all_and", "all_single"
)

plot_group_levels <- c("Control (valid)", "Control (invalid)", "Want", "Modal: allowed/must", "Modal: all")

plot_data <- experimental %>%
  mutate(
    plot_group = case_when(
      domain == "control" & valid_flag == "Y" ~ "Control (valid)",
      domain == "control" & valid_flag == "N" ~ "Control (invalid)",
      domain == "want" ~ "Want",
      domain == "modal" & pair_type %in% c("allowed_and", "allowed_single", "must_and", "must_single") ~ "Modal: allowed/must",
      domain == "modal" & pair_type %in% c("all_and", "all_single") ~ "Modal: all",
      TRUE ~ NA_character_
    ),
    plot_group = factor(plot_group, levels = plot_group_levels),
    plot_pair_type = factor(pair_type, levels = pair_order)
  ) %>%
  filter(!is.na(plot_group))
```

### Proportion of "Follows" Responses (by grouping)

```{r grouped-follows, fig.width=14, fig.height=9}
resp_summary <- plot_data %>%
  mutate(
    cf = if_else(is.na(conclusion_form), "control", as.character(conclusion_form)),
    cf = factor(cf, levels = c("AND", "A", "B", "control"))
  ) %>%
  group_by(plot_group, plot_pair_type, cf, participant) %>%
  summarise(mean_resp = mean(resp, na.rm = TRUE), .groups = "drop") %>%
  group_by(plot_group, plot_pair_type, cf) %>%
  summarise(
    mean = mean(mean_resp, na.rm = TRUE),
    se = sd(mean_resp, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

ggplot(resp_summary, aes(plot_pair_type, mean, fill = cf)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(width = 0.8),
                width = 0.2, linewidth = 0.6) +
  facet_wrap(~ plot_group, scales = "free_x") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_fill_scico_d(palette = "batlow", name = "Conclusion form", na.value = "grey80") +
  labs(
    title = "Proportion of 'Follows' Responses by Group",
    x = "Pair Type",
    y = "Mean 'Follows' Rate"
  ) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, size = 10),
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Underlying descriptives:

```{r grouped-follows-table, results='asis'}
resp_tbl <- resp_summary %>%
  mutate(
    `Mean follows` = scales::percent(mean, accuracy = 0.1),
    `SE` = round(se, 3)
  ) %>%
  select(Group = plot_group, `Pair type` = plot_pair_type, `Conclusion form` = cf,
         `Mean follows`, SE, n)

resp_tbl %>% flextable() %>% autofit()
```

## Paper-ready summary: want_single vs invalid controls

Primary contrast for the paper: the extent to which participants accept the FC-style inference
from a *want + disjunction* premise to a *single-disjunct want* conclusion (`want_single`),
compared to their tendency to accept clearly invalid control inferences.

```{r paper-summary-want-single-vs-invalid, include=FALSE}
paper_primary <- experimental %>%
  mutate(
    paper_group = case_when(
      domain == "want" & pair_type == "want_single" ~ "Want: want_single",
      domain == "control" & valid_flag == "N" ~ "Control: invalid",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(paper_group)) %>%
  group_by(participant, paper_group) %>%
  summarise(
    mean_resp = mean(resp, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  )

paper_primary_wide <- paper_primary %>%
  select(participant, paper_group, mean_resp) %>%
  pivot_wider(names_from = paper_group, values_from = mean_resp)

paper_primary_summary <- paper_primary %>%
  group_by(paper_group) %>%
  summarise(
    n_participants = n_distinct(participant),
    mean = mean(mean_resp, na.rm = TRUE),
    sd = sd(mean_resp, na.rm = TRUE),
    se = sd(mean_resp, na.rm = TRUE) / sqrt(n_participants),
    .groups = "drop"
  ) %>%
  mutate(
    ci_mult = qt(0.975, df = n_participants - 1),
    ci_low = mean - ci_mult * se,
    ci_high = mean + ci_mult * se,
    `Mean follows` = scales::percent(mean, accuracy = 0.1),
    `SD` = scales::percent(sd, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      scales::percent(ci_low, accuracy = 0.1),
      ", ",
      scales::percent(ci_high, accuracy = 0.1),
      "]"
    )
  ) %>%
  select(
    Group = paper_group,
    `Mean follows`,
    SD,
    `95% CI`,
    `N participants` = n_participants
  )

paper_primary_t <- t.test(
  paper_primary_wide$`Want: want_single`,
  paper_primary_wide$`Control: invalid`,
  paired = TRUE
)
paper_primary_diff <- paper_primary_wide$`Want: want_single` - paper_primary_wide$`Control: invalid`
paper_primary_dz <- mean(paper_primary_diff, na.rm = TRUE) / sd(paper_primary_diff, na.rm = TRUE)

paper_primary_text <- glue(
  "Paired t-test on participant means: t({paper_primary_t$parameter}) = {round(paper_primary_t$statistic, 2)}, ",
  "p = {format.pval(paper_primary_t$p.value, digits = 3, eps = .001)}, ",
  "dz = {round(paper_primary_dz, 2)}."
)

list(
  paper_primary_summary = paper_primary_summary,
  paper_primary_text = paper_primary_text
)
```

Primary descriptives:

```{r paper-primary-table, results='asis'}
paper_primary_summary %>% flextable() %>% autofit()
```

Primary comparison test:

`r paper_primary_text`

### Primary comparison (model-based): want_single vs invalid controls

In addition to participant-mean descriptives, we estimate the same contrast using a
mixed-effects logistic regression on trial-level responses.

```{r paper-primary-model, include=FALSE}
paper_primary_model_data <- experimental %>%
  mutate(
    paper_group = case_when(
      domain == "want" & pair_type == "want_single" ~ "Want: want_single",
      domain == "control" & valid_flag == "N" ~ "Control: invalid",
      TRUE ~ NA_character_
    ),
    paper_group = factor(paper_group, levels = c("Control: invalid", "Want: want_single"))
  ) %>%
  filter(!is.na(paper_group)) %>%
  droplevels()

paper_primary_glmer <- glmer(
  resp ~ paper_group + (1 | participant) + (1 | item),
  data = paper_primary_model_data,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

paper_primary_emm_resp <- summary(emmeans(paper_primary_glmer, ~ paper_group, type = "response"))
paper_primary_contrast <- summary(
  contrast(
    emmeans(paper_primary_glmer, ~ paper_group),
    method = list("Want: want_single - Control: invalid" = c(-1, 1))
  ),
  infer = c(TRUE, TRUE)
)

paper_primary_contrast_df <- as.data.frame(paper_primary_contrast) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    odds_ratio = exp(estimate),
    or_low = exp(lcl),
    or_high = exp(ucl)
  )

paper_primary_model_text <- glue(
  "Mixed-effects logistic regression (random intercepts for participant and item): ",
  "OR = {round(paper_primary_contrast_df$odds_ratio[1], 2)}, ",
  "95% CI [{round(paper_primary_contrast_df$or_low[1], 2)}, {round(paper_primary_contrast_df$or_high[1], 2)}], ",
  "z = {round(paper_primary_contrast_df$z.ratio[1], 2)}, ",
  "p = {format.pval(paper_primary_contrast_df$p.value[1], digits = 3, eps = .001)}."
)

list(
  paper_primary_emm_resp = paper_primary_emm_resp,
  paper_primary_contrast = paper_primary_contrast,
  paper_primary_model_text = paper_primary_model_text
)
```

Model-based estimated “follows” rates:

```{r paper-primary-emm-table, results='asis'}
as.data.frame(paper_primary_emm_resp) %>% flextable() %>% autofit()
```

Model-based comparison:

`r paper_primary_model_text`

### Paper-ready summary: all conditions (collapsed over A/B where applicable)

The table below collapses across `A`/`B` conclusion variants for single-disjunct conditions.

```{r paper-summary-all-conditions, include=FALSE}
paper_conditions <- experimental %>%
  mutate(
    paper_condition = case_when(
      domain == "control" & valid_flag == "Y" ~ "Control: valid",
      domain == "control" & valid_flag == "N" ~ "Control: invalid",
      domain == "want" ~ paste0("Want: ", as.character(pair_type)),
      domain == "modal" ~ paste0("Modal: ", as.character(pair_type)),
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(paper_condition)) %>%
  group_by(participant, paper_condition) %>%
  summarise(mean_resp = mean(resp, na.rm = TRUE), .groups = "drop")

paper_conditions_summary <- paper_conditions %>%
  group_by(paper_condition) %>%
  summarise(
    n_participants = n_distinct(participant),
    mean = mean(mean_resp, na.rm = TRUE),
    sd = sd(mean_resp, na.rm = TRUE),
    se = sd(mean_resp, na.rm = TRUE) / sqrt(n_participants),
    .groups = "drop"
  ) %>%
  mutate(
    ci_mult = qt(0.975, df = n_participants - 1),
    ci_low = mean - ci_mult * se,
    ci_high = mean + ci_mult * se,
    `Mean follows` = scales::percent(mean, accuracy = 0.1),
    `SD` = scales::percent(sd, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      scales::percent(ci_low, accuracy = 0.1),
      ", ",
      scales::percent(ci_high, accuracy = 0.1),
      "]"
    )
  ) %>%
  select(
    Condition = paper_condition,
    `Mean follows`,
    SD,
    `95% CI`,
    `N participants` = n_participants
  )

list(paper_conditions_summary = paper_conditions_summary)
```

Condition-level descriptives (participant means):

```{r paper-conditions-table, results='asis'}
paper_conditions_summary %>% flextable() %>% autofit()
```

## Additional comparisons: benchmark modal FC vs want FC

The main text currently reports the primary contrast (`want_single` vs invalid controls). For interpretability, it is also useful to compare FC endorsement rates in *benchmark permission modals* (allowed) to FC endorsement rates under *want*.

We focus on the four directly comparable conditions:

- **Modal (allowed):** `allowed_single` and `allowed_and`
- **Want:** `want_single` and `want_and`

### Descriptives (participant means)

```{r fc_modal_vs_want_descriptives, include=FALSE}
fc_comp <- experimental %>%
  filter(
    (domain == "modal" & pair_type %in% c("allowed_single", "allowed_and")) |
      (domain == "want" & pair_type %in% c("want_single", "want_and"))
  ) %>%
  mutate(
    fc_domain = case_when(
      domain == "modal" ~ "Modal: allowed",
      domain == "want" ~ "Want",
      TRUE ~ NA_character_
    ),
    fc_conclusion = case_when(
      pair_type %in% c("allowed_single", "want_single") ~ "single",
      pair_type %in% c("allowed_and", "want_and") ~ "and",
      TRUE ~ NA_character_
    ),
    fc_domain = factor(fc_domain, levels = c("Modal: allowed", "Want")),
    fc_conclusion = factor(fc_conclusion, levels = c("single", "and"))
  ) %>%
  filter(!is.na(fc_domain), !is.na(fc_conclusion))

fc_subj_means <- fc_comp %>%
  group_by(participant, fc_domain, fc_conclusion) %>%
  summarise(mean_resp = mean(resp, na.rm = TRUE), .groups = "drop")

fc_desc <- fc_subj_means %>%
  group_by(fc_domain, fc_conclusion) %>%
  summarise(
    n_participants = n_distinct(participant),
    mean = mean(mean_resp, na.rm = TRUE),
    sd = sd(mean_resp, na.rm = TRUE),
    se = sd / sqrt(n_participants),
    .groups = "drop"
  ) %>%
  mutate(
    ci_mult = qt(0.975, df = n_participants - 1),
    ci_low = mean - ci_mult * se,
    ci_high = mean + ci_mult * se,
    `Mean follows` = scales::percent(mean, accuracy = 0.1),
    `SD` = scales::percent(sd, accuracy = 0.1),
    `95% CI` = paste0(
      "[",
      scales::percent(ci_low, accuracy = 0.1),
      ", ",
      scales::percent(ci_high, accuracy = 0.1),
      "]"
    )
  ) %>%
  select(
    Domain = fc_domain,
    Conclusion = fc_conclusion,
    `Mean follows`,
    SD,
    `95% CI`,
    `N participants` = n_participants
  )

NULL
```

```{r fc_modal_vs_want_descriptives_table, results='asis'}
fc_desc %>% flextable() %>% autofit()
```

### Model-based comparison (mixed-effects logistic regression)

```{r fc_modal_vs_want_model, include=FALSE}
fc_model <- glmer(
  resp ~ fc_domain * fc_conclusion + (1 | participant) + (1 | item),
  data = fc_comp,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

fc_emm <- emmeans(fc_model, ~ fc_domain * fc_conclusion, type = "response")

# Key contrasts for the paper (on the logit scale, reported as OR):
fc_contrast_domain_within_conclusion <- summary(
  pairs(emmeans(fc_model, ~ fc_domain | fc_conclusion)),
  infer = c(TRUE, TRUE)
)

fc_contrast_conclusion_within_domain <- summary(
  pairs(emmeans(fc_model, ~ fc_conclusion | fc_domain)),
  infer = c(TRUE, TRUE)
)

list(
  fc_model = fc_model,
  fc_emm = fc_emm,
  fc_contrast_domain_within_conclusion = fc_contrast_domain_within_conclusion,
  fc_contrast_conclusion_within_domain = fc_contrast_conclusion_within_domain
)
```

Model-based predicted “follows” rates:

```{r fc_modal_vs_want_emm_table, results='asis'}
as.data.frame(summary(fc_emm, infer = c(TRUE, TRUE))) %>%
  mutate(across(c(prob, asymp.LCL, asymp.UCL), ~ scales::percent(.x, accuracy = 0.1))) %>%
  select(fc_domain, fc_conclusion, prob, asymp.LCL, asymp.UCL, z.ratio, p.value) %>%
  rename(
    Domain = fc_domain,
    Conclusion = fc_conclusion,
    `Predicted follows` = prob,
    `95% CI (low)` = asymp.LCL,
    `95% CI (high)` = asymp.UCL,
    z = z.ratio,
    p = p.value
  ) %>%
  flextable() %>%
  autofit()
```

Domain contrasts (Want vs Modal: allowed), separately for `single` vs `and` conclusions:

```{r fc_modal_vs_want_contrast_domain, results='asis'}
as.data.frame(fc_contrast_domain_within_conclusion) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(round(OR, 2), " [", round(OR_low, 2), ", ", round(OR_high, 2), "]"),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(
    `Conclusion type` = fc_conclusion,
    Contrast = contrast,
    `OR (95% CI)`,
    z,
    p
  ) %>%
  flextable() %>%
  autofit()
```

Conclusion-type contrasts (`and` vs `single`), separately within each domain:

```{r fc_modal_vs_want_contrast_conclusion, results='asis'}
as.data.frame(fc_contrast_conclusion_within_domain) %>%
  mutate(
    lcl = if ("lower.CL" %in% names(.)) lower.CL else asymp.LCL,
    ucl = if ("upper.CL" %in% names(.)) upper.CL else asymp.UCL,
    OR = exp(estimate),
    OR_low = exp(lcl),
    OR_high = exp(ucl),
    `OR (95% CI)` = paste0(round(OR, 2), " [", round(OR_low, 2), ", ", round(OR_high, 2), "]"),
    z = round(z.ratio, 2),
    p = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(
    Domain = fc_domain,
    Contrast = contrast,
    `OR (95% CI)`,
    z,
    p
  ) %>%
  flextable() %>%
  autofit()
```

### Controls: Accuracy (Known Validity Only)

```{r control-accuracy}
controls_only <- experimental %>%
  filter(domain == "control", !is.na(valid_flag))

acc_summary <- controls_only %>%
  group_by(pair_type, valid_flag, participant) %>%
  summarise(mean_acc = mean(accuracy, na.rm = TRUE), .groups = "drop") %>%
  group_by(pair_type, valid_flag) %>%
  summarise(
    mean = mean(mean_acc, na.rm = TRUE),
    se = sd(mean_acc, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

ggplot(acc_summary, aes(valid_flag, mean, fill = pair_type)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(width = 0.8),
                width = 0.2, linewidth = 0.6) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Accuracy on Control Trials",
    x = "Ground Truth",
    y = "Mean Accuracy",
    fill = "Control Type"
  )
```

# Response Time (Experimental Trials)

Trim obvious outliers (e.g., <200 ms or >20 s) and reuse the grouping from above.

```{r rt, fig.width=14, fig.height=9}
rt_trimmed <- experimental %>%
  filter(rt >= 200, rt <= 20000)

rt_summary <- rt_trimmed %>%
  mutate(
    plot_group = case_when(
      domain == "control" & valid_flag == "Y" ~ "Control (valid)",
      domain == "control" & valid_flag == "N" ~ "Control (invalid)",
      domain == "want" ~ "Want",
      domain == "modal" & pair_type %in% c("allowed_and", "allowed_single", "must_and", "must_single") ~ "Modal: allowed/must",
      domain == "modal" & pair_type %in% c("all_and", "all_single") ~ "Modal: all",
      TRUE ~ NA_character_
    ),
    plot_group = factor(plot_group, levels = plot_group_levels),
    plot_pair_type = factor(pair_type, levels = pair_order)
  ) %>%
  filter(!is.na(plot_group)) %>%
  group_by(plot_group, plot_pair_type, participant) %>%
  summarise(mean_rt = mean(rt_sec, na.rm = TRUE), .groups = "drop") %>%
  group_by(plot_group, plot_pair_type) %>%
  summarise(
    mean = mean(mean_rt, na.rm = TRUE),
    se = sd(mean_rt, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

ggplot(rt_summary, aes(plot_pair_type, mean, fill = plot_group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(width = 0.8),
                width = 0.2, linewidth = 0.6) +
  facet_wrap(~ plot_group, scales = "free_x") +
  scale_fill_scico_d(palette = "batlow", name = "Group") +
  labs(
    title = "Mean Response Times (s) by Group",
    x = "Pair Type",
    y = "Response Time (s)",
    fill = "Group"
  ) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, size = 10),
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Underlying descriptives:

```{r rt-table, results='asis'}
rt_tbl <- rt_summary %>%
  mutate(
    `Mean RT (s)` = round(mean, 2),
    `SE` = round(se, 2)
  ) %>%
  select(Group = plot_group, `Pair type` = plot_pair_type, `Mean RT (s)`, SE, n)

rt_tbl %>% flextable() %>% autofit()
```

# Modeling

## Proportion of "Follows" Responses (by grouping)

```{r}
critical <- experimental %>% filter(domain %in% c("modal", "want"))
glmer_critical <- glmer(
  resp ~ domain * pair_type + (1 | participant) + (1| item),
  data = critical,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
  )

tab_model(glmer_critical)
```

```{r}
glmer.emm <- emmeans(glmer_critical, ~ pair_type | domain, type = "response")
format_pairs <- function(df) {
  out <- df %>% mutate(across(where(is.numeric), ~ round(.x, 3)))
  p_col <- intersect(c("adj.p.value", "p.value"), names(out))
  if (length(p_col)) {
    col <- p_col[1]
    out <- out %>%
      mutate(
        !!col := ifelse(.data[[col]] < 0.001, "< 0.001", scales::number(.data[[col]], accuracy = 0.001)),
        .bold_p = ifelse(.data[[col]] != "< 0.001" & suppressWarnings(as.numeric(.data[[col]])) < 0.05, TRUE,
                         ifelse(.data[[col]] == "< 0.001", TRUE, FALSE))
      )
  } else {
    out$.bold_p <- FALSE
  }
  out
}

emm_pairs_glmer <- format_pairs(tidy(pairs(glmer.emm)))
ft_glmer <- flextable(emm_pairs_glmer)
if (".bold_p" %in% names(emm_pairs_glmer)) {
  bold_rows <- which(emm_pairs_glmer$.bold_p %in% TRUE)
  pcol <- intersect(c("adj.p.value", "p.value"), names(emm_pairs_glmer))
  if (length(pcol)) {
    ft_glmer <- bold(ft_glmer, i = bold_rows, j = pcol, bold = TRUE)
  }
  ft_glmer <- delete_columns(ft_glmer, ".bold_p")
}
ft_glmer %>% autofit()
plot(glmer.emm)
```

## Response Time (Experimental Trials)

```{r}
critical <- experimental %>% filter(domain %in% c("modal", "want"))
lmer_critical <- lmer(
  rt ~ domain * pair_type + (1 | participant) + (1| item),
  data = critical,
  control = lmerControl(optimizer = "bobyqa")
  )
tab_model(lmer_critical)
```

```{r}
lmer.emm <- emmeans(lmer_critical, ~ pair_type | domain, type = "response")
emm_pairs_lmer <- format_pairs(tidy(pairs(lmer.emm)))
ft_lmer <- flextable(emm_pairs_lmer)
if (".bold_p" %in% names(emm_pairs_lmer)) {
  bold_rows <- which(emm_pairs_lmer$.bold_p %in% TRUE)
  pcol <- intersect(c("adj.p.value", "p.value"), names(emm_pairs_lmer))
  if (length(pcol)) {
    ft_lmer <- bold(ft_lmer, i = bold_rows, j = pcol, bold = TRUE)
  }
  ft_lmer <- delete_columns(ft_lmer, ".bold_p")
}
ft_lmer %>% autofit()
plot(lmer.emm)
```

# Per-Item "Follows" Rates

```{r per-item-follows}
per_item_summary <- experimental %>%
  filter(domain != "control") %>%
  group_by(item, domain, pair_type, conclusion_form) %>%
  summarise(
    Premise = first(premise),
    Conclusion = first(conclusion),
    follows_rate = mean(resp, na.rm = TRUE),
    n = sum(!is.na(resp)),
    .groups = "drop"
  ) %>%
  mutate(
    Conclusion_form = if_else(is.na(conclusion_form) | conclusion_form == "", "control/NA", as.character(conclusion_form))
  ) %>%
  select(
    Item = item,
    Premise,
    Conclusion,
    Domain = domain,
    Condition = pair_type,
    Conclusion_form,
    follows_rate,
    `N responses` = n
  )

per_item_display <- per_item_summary %>%
  mutate(`Follows %` = scales::percent(follows_rate, accuracy = 0.1)) %>%
  select(-follows_rate)
```

```{r per-item-follows-table, results='asis'}
per_item_display %>% flextable() %>% autofit()
```

```{r per-item-follows-export}
per_item_export_path <- "per_item_follows.xlsx"
write_xlsx(per_item_summary, per_item_export_path)
glue("Per-item follows proportions exported to: {per_item_export_path}")
```
